---
title: 'Creating a prior distribution for audit sampling'
author: Koen Derks
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{Creating prior distributions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteDepends{jfa}
  %\VignetteKeywords{audit, evaluation, planning, prior}
  %\VignettePackage{jfa}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(jfa)
```

## Introduction

This vignette discusses how to set up a prior distribution for audit sampling
using the `auditPrior()` function in the **jfa** package.

Bayesian statistics enables auditors to incorporate existing information about
the misstatement into the sampling procedure via the prior distribution. After
observing data, the information in the prior is updated to come to an overall
conclusion about the misstatement in the population. The advantage of using a
prior distribution is that it makes the underlying assumptions explicit and can
therefore decrease the amount of audit work required to achieve the desired
assurance. For example, when you have information from the auditee's internal
controls that indicates a low-risk profile, this information can be incorporated
into the prior distribution in order to require less evidence from tests of
details. Note that all incorporated prior information should be justified with
appropriate audit evidence. What information can be incorporated into the prior
distribution depends on what type of information is available, the quality of
that information, and the situation at hand.

## Creating a prior distribution

The `auditPrior()` function is used to specify a prior distribution for audit
sampling. Below is an enumeration of the several ways that a prior distribution
can be constructed using this function.

### Default priors (`method = 'default'`)

The default prior distributions are created using `method = 'default'`.
**jfa**'s default priors satisfy two criteria: 1) they contain relatively little
information about the population misstatement and 2) they are proper (i.e., they
integrate to 1). The default priors in **jfa** are:

* `likelihood = 'poisson'`: gamma($\alpha$ = 1, $\beta$ = 1)
* `likelihood = 'binomial'`: beta($\alpha$ = 1, $\beta$ = 1)
* `likelihood = 'hypergeometric'`: beta-binomial(N, $\alpha$ = 1, $\beta$ = 1)
* `likelihood = 'normal'`: normal($\mu$ = 0, $\sigma$ = 1000)
* `likelihood = 'uniform'`: uniform(min = 0, max = 1) 
* `likelihood = 'cauchy'`: Cauchy($x_0$ = 0, $\gamma$ = 1000)
* `likelihood = 't'`: Student-t(df = 1)
* `likelihood = 'chisq'`: chi-squared(df = 1)

```{r}
prior <- auditPrior(method = "default", likelihood = "binomial")
summary(prior)
```

All prior distributions can be visually inspected via the `plot()` function.

```{r fig.align="center", fig.height=4, fig.width=6}
plot(prior)
```

### Priors with custom parameters (`method = 'param'`)

You can manually specify the parameters of the prior distribution with
`method = 'param'` and the `alpha` and `beta` arguments, which correspond to the
first and (optionally) second parameter of the prior as described above. For
example, the commands below create a beta(2, 10) prior distribution, a
normal(0.025, 0.05) prior distribution, and a Student-t(0.01) prior
distribution.

```{r}
auditPrior(method = "param", likelihood = "binomial", alpha = 2, beta = 10)
auditPrior(method = "param", likelihood = "normal", alpha = 0.025, beta = 0.05)
auditPrior(method = "param", likelihood = "t", alpha = 0.01)
```

### Improper priors (`method = 'strict'`)

You can construct an improper prior distribution with classical properties using
`method = 'strict'`. The posterior distribution of from this prior yields the
same results as the classical methodology with respect to sample sizes and upper
limits, but is only proper once a single non-misstated unit is present in the 
sample (Derks et al., 2022). For example, the command below creates an improper
beta(1, 0) prior distribution.

*This method requires the `poisson`, `binomial`, or `hypergeometric` likelihood*

```{r}
auditPrior(method = "strict", likelihood = "binomial")
```

### Impartial priors (`method = 'impartial'`)

You can incorporate the assumption that tolerable misstatement is equally likely
as intolerable misstatement (Derks et al., 2022) using `method = 'impartial'`.
For example, the command below creates an impartial beta prior distribution for
a performance materiality of 5 percent.

*This method requires that you specify a value for the `materiality`.*

*This method requires the `poisson`, `binomial`, or `hypergeometric` likelihood*

```{r}
auditPrior(method = "impartial", likelihood = "binomial", materiality = 0.05)
```

### Priors with custom prior probabilities (`method = 'hyp'`)

You can manually assign prior probabilities to the hypothesis of tolerable 
misstatement and the hypotheses of intolerable misstatement (Derks et al., 2021)
with `method = 'hyp'` in combination with `p.hmin`. For example, the command
below incorporates the information that the hypothesis of tolerable misstatement 
has a probability of 60% into a beta prior distribution.

*This method requires that you specify a value for the `materiality`.*

*This method requires the `poisson`, `binomial`, or `hypergeometric` likelihood*

```{r}
auditPrior(method = "hyp", likelihood = "binomial", materiality = 0.05, p.hmin = 0.6)
```

### Priors based on the Audit Risk Model (`method = 'arm'`)

You can translate risk assessments from the Audit Risk Model (inherent risk and
internal control risk) into a prior distribution (Derks et al., 2021) using
`method = 'arm'` in combination with the `ir` and `cr` arguments. For example,
the command below incorporates the information that the inherent risk is equal
to 90% and internal control risk is equal to 60% into a beta prior distribution.

*This method requires the `poisson`, `binomial`, or `hypergeometric` likelihood*

```{r}
auditPrior(method = "arm", likelihood = "binomial", materiality = 0.05, ir = 0.9, cr = 0.6)
```

### Priors based on the Bayesian Risk Assessment Model (`method = 'bram'`)

You can incorporate information about the mode and the upper bound of the prior
distribution using `method = 'bram'`. For example, the code below incorporates
the information that the mode of the prior distribution is 1% and the upper
bound is 60% into a beta prior distribution.

*This method requires the `poisson`, `binomial`, or `hypergeometric` likelihood*

```{r}
auditPrior(method = "bram", likelihood = "binomial", expected = 0.01, materiality = 0.05, ub = 0.6)
```

### Priors based on an earlier sample (`method = 'sample'`)

You can incorporate information from an earlier sample into the prior
distribution (Derks et al., 2021) using `method = 'sample'` in combination with
`x` and `n`. For example, the command below incorporates the information from
an earlier sample of 30 items in which 0 misstatements were found into a beta
prior distribution.

*This method requires the `poisson`, `binomial`, or `hypergeometric` likelihood*

```{r}
auditPrior(method = "sample", likelihood = "binomial", x = 0, n = 30)
```

### Priors based on a weighted earlier sample (`method = 'factor'`)

You can incorporate information from last years results, weighted by a factor
(Derks et al., 2021), into the prior distribution using `method = 'factor'` in
combination with `x` and `n`. For example, the command below incorporates the
information from a last years results (a sample of 58 items in which 0
misstatements were found), weighted by a factor 0.7, into a beta prior
distribution.

*This method requires the `poisson`, `binomial`, or `hypergeometric` likelihood*

```{r}
auditPrior(method = "factor", likelihood = "binomial", x = 0, n = 58, factor = 0.7)
```

## Using a prior distribution

The objects returned by the `auditPrior()` function can be used as input for the
`prior` argument in the `planning()` and `evaluation()` functions. Below is
small demonstration of how the prior distribution can be constructed using these
functions.

### Planning a sample

The prior distribution can be used the planning stage to calculate a minimum
sample size by providing the objected returned by the `auditPrior()` function to
the `planning()` function. For example, the command below calculates the minimum
sample size to test the misstatement in the population against a performance
materiality of 5%, while using a beta(1, 10) prior distribution.

```{r}
prior <- auditPrior(method = "param", likelihood = "binomial", alpha = 1, beta = 10)
planning(materiality = 0.05, likelihood = "binomial", prior = prior)
```

### Evaluating a sample

The prior distribution can be used the evaluation stage by providing the
objected returned by the `auditPrior()` function to the `evaluation()` function.
For example, the command below evaluates the misstatement in the population with
respect to a performance materiality of 5% after seeing a sample of 60 items
with 1 misstatement, while using a normal(0.025, 0.05) prior distribution.

```{r}
prior <- auditPrior(method = "param", likelihood = "normal", alpha = 0.025, beta = 0.05)
eval <- evaluation(materiality = 0.05, x = 1, n = 60, prior = prior)
summary(eval)
plot(eval)
```

## References

- Derks, K., de Swart, J., Wagenmakers, E.-J., and Wetzels, R. (2022). An impartial Bayesian hypothesis test for audit sampling. *PsyArXiv*. https://doi.org/10.31234/osf.io/8nf3e
- Derks, K., de Swart, J., van Batenburg, P., Wagenmakers, E.-J., and Wetzels, R. (2021). Priors in a Bayesian audit: How integration of existing information  into the prior distribution can improve audit transparency and efficiency. *International Journal of Auditing*, 25(3), 621-636. https://doi.org/10.1111/ijau.12240
