---
title: "Data auditing: Get started"
author: Koen Derks
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
bibliography: references.bib
csl: apa.csl
vignette: >
  %\VignetteIndexEntry{Data auditing: Get started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteDepends{jfa}
  %\VignetteKeywords{digit, tests, Bayesian}
  %\VignettePackage{jfa}
  %\VignetteEncoding{UTF-8}
nocite: |
  @cinelli_2018, @joenssen_2015, @stephane_2020
---

```{r, include = FALSE}
library(jfa)
```

## Introduction

Welcome to the 'Data auditing' vignette of the **jfa** package. Here you can
find a simple explanation of the functions in the package that facilitate data
auditing. For more detailed explanations of each function, read the other
vignettes on the [package website](https://koenderks.github.io/jfa/).

## Functions and intended usage

Below you can find an explanation of the available data auditing functions in
**jfa**.

- [`digit_test()`](#test-digit-distributions-with-digit-test)
- [`repeated_test()`](#test-for-repeated-values-with-repeated-test)

### Test digit distributions with `digit_test()`

[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)

The function `digit_test()` takes a vector of numeric values, extract the
requested digits, and compares the frequencies of these digits to a reference
distribution. By default, the function performs a frequentist hypothesis test of
the null hypothesis that the digits are distributed according to the reference
distribution and produces a *p* value. When a prior is specified, the function
performs a Bayesian hypothesis test of the null hypothesis that the digits are
distributed according to the reference distribution against the alternative
hypothesis that the digits are not distributed according to the reference
distribution and produces a Bayes factor [@kass_1995].

*Full function with default arguments:*

```r
digit_test(
  x,
  check = c("first", "last", "firsttwo"),
  reference = "benford",
  prior = FALSE
)
```

*Supported options for the `check` argument:*

| `check` | Returns |
| :----------- | :----------- |
| `fist` | First digit |
| `firsttwo` | First and second digit |
| `last` | Last digit |

*Supported options for the `reference` argument:*

| `check` | Returns |
| :----------- | :----------- |
| `benford` | Benford's law |
| `uniform` | Uniform distribution |
| Vector of probabilities | Custom distribution |

*Example usage:*

Benford's law [@benford_1938] is a principle that describes a pattern in many
naturally-occurring numbers. According to Benford's law, each possible leading
digit *d* in a naturally occurring, or non-manipulated, set of numbers occurs
with a probability:

<img src="https://latex.codecogs.com/svg.image?p(d_i)&space;=&space;\text{log}_{10}(1&space;+&space;\frac{1}{d_i})" title="p(d_i) = \text{log}_{10}(\frac{1}{d_i})" />

The distribution of leading digits in a data set of financial transaction values
(e.g., the `sinoForest` data) can be extracted and tested against the expected
frequencies under Benford's law using the code below.

```{r}
# Frequentist hypothesis test
digit_test(sinoForest$value, check = "first", reference = "benford")

# Bayesian hypothesis test using default prior
digit_test(sinoForest$value, check = "first", reference = "benford", prior = TRUE)
```

### Test for repeated values with `repeated_test()`

[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)

The function `repeated_test()` analyzes the frequency with which values get
repeated within a set of numbers. Unlike Benford's law, and its generalizations,
this approach examines the entire number at once, not only the first or last
digit. For the technical details of this procedure, see [@simonsohn_2019].

*Full function with default arguments:*

```r
repeated_test(
  x,
  check = "last",
  method = "af",
  samples = 2000
)
```

*Supported options for the `check` argument:*

| `check` | Returns |
| :----------- | :----------- |
| `last` | Last decimal |
| `lasttwo` | Last two decimals |
| `all` | All decimals |

*Supported options for the `method` argument:*

| `check` | Returns |
| :----------- | :----------- |
| `af` | Average frequency |
| `entropy` | Entropy |

*Example usage:*

In this example, we analyze a data set from a (retracted) paper that describes
three experiments run in Chinese factories, where workers were nudged to use
more hand-sanitizer. These data were shown to exhibited two classic markers of
data tampering: impossibly similar means and the uneven distribution of last
digits [@yu_2018]. We can use the `rv.test()` function to test if these data
also contain a greater amount of repeated values than expected if the data were
not tampered with.

```{r}
repeated_test(sanitizer$value, check = "lasttwo", samples = 5000)
```

## Benchmarks

To validate the statistical results, **jfa**'s automated
[unit tests](https://github.com/koenderks/jfa/tree/development/tests/testthat)
regularly verify the main output from the package against the following
benchmarks:

- [benford.analysis](https://cran.r-project.org/package=benford.analysis) (R package version 0.1.5)
- [BenfordTests](https://cran.r-project.org/package=BenfordTests) (R package version 1.2.0)
- [BeyondBenford](https://cran.r-project.org/package=BeyondBenford) (R package version 1.4)

## References
<div id="refs"></div>
