---
title: "Get started"
author: Koen Derks
date: "last modified: 29-06-2021"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Get started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteDepends{jfa}
  %\VignetteKeywords{audit, evaluation, jfa, planning, sampling}
  %\VignettePackage{jfa}
  %\VignetteEncoding{UTF-8}
---

Welcome to the 'Get started' page of the `jfa` package. In this vignette you are able to find detailed examples of how you can incorporate the functions provided by the package into your statistical sampling workflow.

## Example data

To concretely illustrate `jfa`'s functionality, we consider the `BuildIt` data set that is included in the package (for more info, see `?BuildIt`). This data set contains a population of 3500 invoices paid to a fictional construction company. Each invoice has an identification number (`ID`), a recorded value (`bookValue`), and a corresponding audit (true) value (`auditValue`).

*Note: The information in the `auditValue` column is added for illustrative purposes, as it is unknown to the auditor before having audited any items from a sample.*

First, we load in the data set and display the first 10 invoices in the population.

```{r}
library(jfa)

data('BuildIt')
head(BuildIt, n = 10)
```

For a walkthrough of `jfa`'s workflow functionality using the `BuildIt` data set, see [The audit sampling workflow](https://koenderks.github.io/jfa/articles/v1_sampling_workflow.html). For a Bayesian version of the the walkthrough, see [The Bayesian audit sampling workflow](https://koenderks.github.io/jfa/articles/v2_bayesian_sampling_workflow.html).

## (Optional) Using `auditPrior()`: The basics

The `auditPrior()` function allows you to create a prior distribution for the misstatement in the population that can be used to incorporate existing information into the sampling workflow. Prior distributions can be constructed based on various types of audit information. See the vignette [Prior distributions](https://koenderks.github.io/jfa/articles/v3_prior_distributions.html) for a detailed explanation of the types of audit information that `jfa` is able to incorporate into a prior distribution. Bayesian planning and evaluation using the prior distribution can be performed by specifying the returned object from the `auditPrior()` function as input for the `prior` argument in the `planning()` and `evaluation()` functions.

## Using `planning()`: The basics

Planning a sample using the `planning()` function requires knowledge of the objective of the sampling procedure, the assumed distribution of the data (`binomial`, `poisson`, or `hypergeometric`), and the expected errors in the sample. However, it is advised to set the value for the expected errors in the sample conservatively to minimize the probability that the observed errors in the sample exceed the expected errors, which would imply that insufficient work has been done in the end.

Because we have access to the recorded values of the invoices in the population, we are going to consider each monetary unit in the population as a possible unit of inference. The audit standards assume that such data are distributed according to the Poisson distribution and so we will assume this distribution in this example as well. For illustrative purposes, we do not expect to find any errors in the sample.

### Testing against a performance materiality

First, we take a look at how you can use the `planning()` function to construct a sample with the objective of testing the misstatement in the population against a performance materiality (i.e., the maximum tolerable misstatement in the population). In this example, we will set the performance materiality at 5% of the total value of the population.

Sampling objective: *Calculate a minimal sample size such that, when zero misstatements are found in the sample, there is a 95% 'chance' that the misstatement in the population is lower than 5% of the population value.*

Constructing a sample with this objective in mind can be done using the code below (specifically by specifying the `materiality` argument). A summary of the output can be obtained from the `summary()` function. As you can see, the minimal sample size to achieve 95% assurance with respect to the performance materiality of 5% is 60 monetary units.

```{r}
stage1 <- planning(materiality = 0.05, expectedError = 0, likelihood = 'poisson', confidence = 0.95, N = 3500)
summary(stage1)
```

### Obtaining a minimum precision

Next, we take a look at how you can use the `planning()` function to construct a sample with the objective of obtaining a minimum precision. The precision is an indication of the accuracy of your estimate of the misstatement, as it is defined as the difference between the most likely misstatement and the upper confidence bound on the misstatement. For this example, we will set the minimum precision to 2% of the population value.

Sampling objective: *Calculate a minimal sample size such that, when zero misstatements are found in the sample, there is a 95% 'chance' that the misstatement in the population is at most 2% above the most likely misstatement.*

Planning a sample with this objective can be done using the code below (specifically by specifying the `minPrecision` argument). As you can see, the minimal sample size for to achieve a precision of at least 2% is 150 monetary units.

```{r}
stage1 <- planning(minPrecision = 0.02, expectedError = 0, likelihood = 'poisson', confidence = 0.95, N = 3500)
summary(stage1)
```

## Using `selection()`: The basics

Selecting a sample using the `selection()` function requires knowledge of the sampling units (i.e., units of inference) in the population. Items can be selected from the population using *record sampling* (also known as attribute sampling) using `units = 'records'`, or using *monetary unit sampling (MUS)* using `units = 'mus'`. Selection also requires knowledge of the sampling algorithm. Sampling units can be selected with a *random sampling* scheme using `algorithm = 'random'`, with a *cell sampling* scheme using `algorithm = 'cell'`, or with a *fixed interval sampling* (also known as systematic sampling) scheme using `algorithm = 'interval'`.

### Record sampling

First, we take a look at how you can use the `selection()` function to perform random sampling from the items in the population. As an example, the code below samples 60 invoices from the `BuildIt` data set using a *random record sampling* scheme.

```{r}
stage2 <- selection(population = BuildIt, sampleSize = 60, units = 'records', algorithm = 'random')
summary(stage2)
```

### Monetary unit sampling (MUS)

Next, we take a look at how you can use the `selection()` function to perform fixed interval sampling using the monetary units in the population as sampling units. As an example, the code below samples 150 monetary units from the `BuildIt` data set using a *fixed interval monetary unit sampling* scheme.

```{r}
stage2 <- selection(population = BuildIt, sampleSize = 150, units = 'mus', algorithm = 'interval', bookValues = 'bookValue')
summary(stage2)
```

### Extracting the sample

The selected sample is stored in the object that is returned by the `selection()` function. It can be accesses or extracted by indexing it via `$sample`. Let's take a look at the first ten rows of the previously selected sample of 60 invoices.

```{r}
stage2 <- selection(population = BuildIt, sampleSize = 60, units = 'records', algorithm = 'random')

sample <- stage2$sample
head(sample, n = 10)
```

## Using `evaluation()`: The basics

After saving the sample and annotating the invoices in the sample with their audit values, you can evaluate the misstatement in the sample with the `evaluation()` function. The function can also be used with summary statistics from a data sample. For a more elaborate explanation of how to use this function, see the package vignettes [Testing misstatement](https://koenderks.github.io/jfa/articles/v4_testing_misstatement.html) and [Estimating misstatement](https://koenderks.github.io/jfa/articles/v5_estimating_misstatement.html).

### Summary statistics from the sample

First, let's take a look at how you can use the `evaluation()` function to evaluate the misstatement in the population using summary statistics from a sample. Suppose that in the previously selected sample of 60 invoices you have found that 1 invoice is missing. Using `nSumstats = 60` and `kSumstats = 1` you can provide these outcomes of the sample to the `evaluation()` function. Don't forget to specify your sampling objectives using the `materiality` or `minPrecision` arguments. In this example, a performance materiality of 5% applies.

Sampling objective: *Evaluate, on the basis of summary statistics of a sample, whether the misstatement in the population exceeds the allocated performance materiality such that there is a 5% 'chance' of incorrectly concluding that the population is free of material misstatement.*

```{r}
stage4 <- evaluation(materiality = 0.05, method = 'binomial', confidence = 0.95, N = 3500, nSumstats = 60, kSumstats = 1)
summary(stage4)
```

As you can see, the upper bound on the misstatement is higher than 5% and thus the sample does not provide sufficient evidence to conclude that the misstatement is lower than 5%.

### Annotated sample

Next, we take a look at how you can use the `evaluation()` function to evaluate the misstatement using an annotated sample. Suppose that you have audited the 60 invoices in the sample and have found 1 misstatement. 

```{r}
sample$auditValue    <- sample$bookValue
sample$auditValue[1] <- sample$auditValue[1] - 100
```

You can evaluate the misstatement in the annotated sample using the `sample`, `bookValues`, `auditValues`, and `counts` arguments. For example, the code below evaluates the misstatement in the population with respect to the performance materiality of 5% using the commonly used Stringer bound. You can find more information about which evaluation methods are implemented on the [home page](https://koenderks.github.io/jfa/).

Sampling objective: *Evaluate, on the basis of an annotated sample, whether the misstatement in the population exceeds the allocated performance materiality such that there is a 5% 'chance' of incorrectly concluding that the population is free of material misstatement.*

```{r}
stage4 <- evaluation(materiality = 0.05, method = 'stringer', confidence = 0.95,
                     sample = sample, bookValues = 'bookValue', auditValues = 'auditValue',
                     counts = sample$count, N = 3500)
summary(stage4)
```

## Using `report()`: The basics

After obtaining the result from the `evaluation()` function you can generate a report containing the data, the statistical results and their interpretation, and the conclusion of the sampling procedure with respect to the input for `materiality` and `minPrecision`. The report can be automatically generated by providing the object returned by the `evaluation()` function to the `report()` function.

```{r, eval = FALSE}
stage4 <- evaluation(materiality = 0.05, method = 'stringer', confidence = 0.95,
                     sample = sample, bookValues = 'bookValue', auditValues = 'auditValue',
                     counts = sample$count, N = 3500)

report(stage4, file = 'report.html', format = 'html_document') # Generates .html report
```