---
title: "Testing the misstatement"
author: Koen Derks
date: "last modified: 19-11-2020"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Testing the misstatement}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteDepends{jfa}
  %\VignetteKeywords{audit, evaluation, jfa, planning, prior}
  %\VignettePackage{jfa}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height=4, fig.width=6)
library(jfa)
```

## Testing in an audit sampling context

In an audit sampling test the auditor generally assigns performance materiality, $\theta_{max}$, to the population which expresses the maximum tolerable misstatement (as a fraction or a monetary amount). The auditor then inspects a sample of the population to compare the following two hypotheses:

$$H_-:\theta<\theta_{max}\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\, H_+:\theta\geq\theta_{max}$$. 

The `evaluation()` function allows you to make a statement about the credibility of these two hypotheses after inspecting a sample. The output for testing as discussed in this vignette is only displayed when you enter a value for `materiality` argument.

## Frequentist hypothesis testing using the *p*-value

This will be added in a future version of `jfa`.

## Bayesian hypothesis testing using the Bayes factor

Bayesian hypothesis testing uses the Bayes factor, $BF_{-+}$ or $BF_{+-}$, to make a statement about the evidence provided by the sample in support for one of the two hypotheses $H_-$ or $H_+$. The subscript The Bayes factor denotes which hypothesis it favors. By default, the `evaluation()` function returns the value for $BF_{-+}$.

As an example of how to interpret the Bayes factor, the value of $BF_{-+} = 10$ (provided by the `evluation()` function) can be interpreted as: *the data are 10 times more likely to have occurred under the hypothesis $H_-:\theta<\theta_{max}$ than under the hypothesis $H_+:\theta\geq\theta_{max}$*. $BF_{-+} > 1$ indicates evidence for $H_-$, while $BF_{-+} < 1$ indicates evidence for $H_+$. 

| $BF_{-+}$ | Strength of evidence |
|---------|-------|
| $< 0.01$ | Extreme evidence for $H_+$ |
| $0.01 - 0.033$ | Very strong evidence for $H_+$ |
| $0.033 - 0.10$ | Strong evidence for $H_+$ |
| $0.10 - 0.33$ | Moderate evidence for $H_+$ |
| $0.33 - 1$ | Anecdotal evidence for $H_+$ |
| $1$ | No evidence for $H_-$ or $H_+$ |
| $1 - 3$ | Anecdotal evidence for $H_-$ |
| $3 - 10$ | Moderate evidence for $H_-$ |
| $10 - 30$ | Strong evidence for $H_-$ |
| $30 - 100$ | Very strong evidence for $H_-$ |
| $> 100$ | Extreme evidence for $H_-$ |

### Example 

As an example, consider that an auditor wants to verify whether the population contains less than 5 percent misstatement, implying the hypotheses $H_-:\theta<0.05$ and $H_+:\theta\geq0.05$. They have taken a sample of 40 items, of which 1 contained an error. The prior distribution is assumed to be a non-informative $beta(1,1)$ prior. 

The output below shows that $BF_{-+}=30.28$, implying that there is very strong evidence for $H_-$, the hypothesis that the population contains misstatements lower than 5 percent of the population. 

```{r}
prior <- auditPrior(confidence = 0.95, materiality = 0.05, method = "none", likelihood = "binomial")
evaluation(confidence = 0.95, materiality = 0.05, nSumstats = 40, kSumstats = 1, prior = prior)
```

### Sensitivity to the prior distribution

In audit sampling, the Bayes factor is dependent on the prior distribution for $\theta$. As a rule of thumb, when the prior distribution is very uninformative with respect to the misstatement parameter $\theta$, the Bayes factor overestimates the evidence in favor of $H_-$. You can mitigate this dependency using `method = "median"` in the `auditPrior()` function, which constructs a prior distribution that is impartial with respect to the hypotheses $H_-$ and $H_+$.  

The output below shows that $BF_{-+}=3.08$, implying that there is anecdotal evidence for $H_-$, the hypothesis that the population contains misstatements lower than 5 percent of the population. 

```{r}
prior <- auditPrior(confidence = 0.95, materiality = 0.05, method = "median", likelihood = "binomial")
evaluation(confidence = 0.95, materiality = 0.05, nSumstats = 40, kSumstats = 1, prior = prior)
```