---
title: Creating a prior distribution for audit sampling
author: Koen Derks
output: 
  html_document:
    toc: true
    toc_depth: 3
bibliography: references.bib
csl: apa.csl
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(jfa)
```

## Introduction

This vignette will walk you through the process of setting up a prior
distribution for audit sampling using the `auditPrior()` function in the **jfa**
package.

The power of Bayesian statistics lies in its ability to incorporate pre-existing
information about the misstatement into the sampling procedure through the prior
distribution. Once data is observed, the prior information is updated to draw an
overall conclusion about the misstatement in the population. The use of a prior
distribution has the advantage of making the underlying assumptions explicit,
potentially reducing the amount of audit work needed to achieve the desired
assurance. The type of information that can be incorporated into the prior
distribution depends on the availability and quality of the information, as well
as the specifics of the audit.

For example, if you have information from the auditee's internal control systems
indicating well-segregated duties and numerous effective controls in place, this
suggests a relatively low probability of misstatement. This risk-mitigating
information can be incorporated into the prior distribution, thereby lessening
the need for evidence from the sample. It is important to note that all
information integrated into the prior distribution should be substantiated with
appropriate and sufficient audit evidence. Specifically, the auditor should be
able to justifying the translation of audit information into a prior
distribution. This vignette addresses this issue.

## Creating a prior distribution

The `auditPrior()` function is used to define a prior distribution for audit
sampling. The following are various ways this function can be used to construct
a prior distribution.

### Default priors

The default prior distributions in **jfa** are specified using
`method = 'default'`. These default priors meet two criteria: 1) they hold
relatively minimal information about the population misstatement, and 2) they
are proper, meaning they integrate to one.

* `likelihood = 'poisson'`: gamma($\alpha$ = 1, $\beta$ = 1)
* `likelihood = 'binomial'`: beta($\alpha$ = 1, $\beta$ = 1)
* `likelihood = 'hypergeometric'`: beta-binomial(N, $\alpha$ = 1, $\beta$ = 1)
* `likelihood = 'normal'`: normal($\mu$ = 0, $\sigma$ = 1000)
* `likelihood = 'uniform'`: uniform(min = 0, max = 1) 
* `likelihood = 'cauchy'`: Cauchy($x_0$ = 0, $\gamma$ = 1000)
* `likelihood = 't'`: Student-t(df = 1)
* `likelihood = 'chisq'`: chi-squared(df = 1)
* `likelihood = 'exponential'`: exponential($\lambda$ = 1)

```{r}
prior <- auditPrior(method = "default", likelihood = "binomial")
summary(prior)
```

After setting up the prior distribution, you can visually examine it using the
`plot()` function.

```{r fig.align="center", fig.height=4, fig.width=6}
plot(prior)
```

Additionally, the `predict()` function generates the predictions of the prior
distribution at the data level for a sample of $n$ items. For instance, the
command below calls for the prediction of the default beta(1, 1) prior for a
hypothetical sample of $n$ = 6 items.

```{r}
predict(prior, n = 6)
```

You can visualize the predictions of the prior distribution using the `plot()`
function.

```{r fig.align="center", fig.height=4, fig.width=6}
plot(predict(prior, n = 10))
```

### Priors with custom parameters

You can manually set the parameters of the prior distribution using
`method = 'param'` along with the `alpha` and `beta` arguments. These correspond
to the first and (optionally) second parameter of the prior distribution as
shown in the default prior section. For instance, the commands below will create
a beta(2, 10) prior distribution, a normal(0.025, 0.05) prior distribution, and
a Student-t(0.01) prior distribution.

```{r}
auditPrior(method = "param", likelihood = "binomial", alpha = 2, beta = 10)
auditPrior(method = "param", likelihood = "normal", alpha = 0.025, beta = 0.05)
auditPrior(method = "param", likelihood = "t", alpha = 0.01)
```

### Improper priors

An improper prior distribution with classical properties can be constructed
using `method = 'strict'`. The posterior distribution from this prior produces
the same results as the classical methodology in terms of sample sizes and upper
limits. However, it only becomes proper once a single non-misstated item is
present in the sample [@derks_2022]. The command below creates an improper
beta(1, 0) prior distribution.

*Note: This method requires the `poisson`, `binomial`, or `hypergeometric` likelihood.*

```{r}
auditPrior(method = "strict", likelihood = "binomial")
```

### Impartial priors

To incorporate the assumption that tolerable misstatement is equally likely as
intolerable misstatement [@derks_2022], you can use `method = 'impartial'`. For
instance, the command below generates an impartial beta prior distribution for a
performance materiality of five percent.

*Note: This method requires a value for the `materiality` and the `poisson`, `binomial`, or `hypergeometric` likelihood.*

```{r}
auditPrior(method = "impartial", likelihood = "binomial", materiality = 0.05)
```

### Priors based on risk of material misstatement

To manually set prior probabilities for the hypothesis of tolerable misstatement
and the hypotheses of intolerable misstatement [@derks_2021], you can use
`method = 'hyp'` in conjunction with the `p.hmin` argument. For instance, the
command below integrates the information that the hypothesis of intolerable
misstatement has a 40 percent probability into a beta prior distribution. This
is equal to setting the risk of material misstatement (RMM) to 40 percent.

*Note: This method requires a value for the `materiality` and the `poisson`, `binomial`, or `hypergeometric` likelihood.*

```{r}
auditPrior(method = "hyp", likelihood = "binomial", materiality = 0.05, p.hmin = 1 - 0.4)
```

### Priors based on the Audit Risk Model

To convert risk assessments from the Audit Risk Model, including inherent risk
and internal control risk, into a prior distribution [@derks_2021], you can use
`method = 'arm'` along with the `ir` and `cr` arguments. For instance, the
command below integrates the information that the inherent risk is 90 percent
and the internal control risk is 60 percent into a beta prior distribution.

*Note: This method requires the `poisson`, `binomial`, or `hypergeometric` likelihood.*

```{r}
auditPrior(method = "arm", likelihood = "binomial", materiality = 0.05, ir = 0.9, cr = 0.6)
```

### Priors based on the Bayesian Risk Assessment Model

Incorporating information about the most likely value and the upper limit of the 
population misstatement can be done using `method = 'bram'`. For instance, the
following code demonstrates how to embed the information that the most likely
misstatement in the population is one percent and the 95 percent upper limit on
this misstatement is 60 percent into a beta prior distribution.

*Note: This method requires the `poisson`, `binomial`, or `hypergeometric` likelihood.*

```{r}
auditPrior(method = "bram", likelihood = "binomial", expected = 0.01, materiality = 0.05, ub = 0.6)
```

### Priors based on an earlier sample

You can use the `method = 'sample'` along with the `x` and `n` arguments to
integrate information from an earlier sample into the prior distribution
[@derks_2021]. For instance, the code below demonstrates how to include the data
from a previous sample of $n$ = 30 items where $k$ = 0 misstatements were
detected into a beta prior distribution.

*Note: This method requires the `poisson`, `binomial`, or `hypergeometric` likelihood.*

```{r}
auditPrior(method = "sample", likelihood = "binomial", x = 0, n = 30)
```

### Priors based on a weighted earlier sample

You can integrate information from the previous year's results, adjusted by a
factor [@derks_2021], into the prior distribution using `method = 'power'` in
conjunction with the `x` and `n` arguments. For instance, the code below
demonstrates how to include the data from the previous year's results (i.e., a
sample of $n$ = 58 items with $k$ = 0 misstatements found), weighed by a factor
of 0.7, into a beta prior distribution. This factor `delta` of 0.7 could be
derived from the fact that 70 percent of the auditee's contracts are with
businesses that have remained unchanged over the years, implying that last
year's population is for 70 percent comparable to the population of this year.

*Note: This method requires the `poisson`, `binomial`, or `hypergeometric` likelihood.*

```{r}
auditPrior(method = "power", likelihood = "binomial", x = 0, n = 58, delta = 0.7)
```

### Nonparametric prior distributions

You can establish the prior based on samples from the prior distribution using
`method = 'nonparam'` along with the `samples` argument. For instance, the code
below sets up a prior based on 1000 samples from a beta(1, 10) distribution.

*Note: The `likelihood` argument is not required and will be ignored in this method.*

```{r}
auditPrior(method = "nonparam", samples = rbeta(1000, 1, 10))
```

## Using a prior distribution

The objects returned by the `auditPrior()` function can be used as input for the
`prior` argument in both the `planning()` and `evaluation()` functions. Here
follows a brief demonstration of how to construct the prior distribution using
these functions.

### Planning a sample

The prior distribution can be employed during the planning stage to compute a
minimum sample size. This is done by supplying the object returned by the
`auditPrior()` function to the `planning()` function via its `prior` argument.
For instance, the command below calculates the minimum sample size required to
test the misstatement in the population against a performance materiality of
five percent, using a beta(1, 10) prior distribution. The resulting sample size
is $n$ = 49.

```{r}
prior <- auditPrior(method = "param", likelihood = "binomial", alpha = 1, beta = 10)
planning(materiality = 0.05, likelihood = "binomial", prior = prior)
```

### Evaluating a sample

The prior distribution can also be used during the evaluation stage by providing
the object returned by the `auditPrior()` function to the `evaluation()`
function via its `prior` argument. For example, the command below evaluates the
misstatement in the population with respect to a performance materiality of five
percent after observing a sample of $n$ = 60 items with $k$ = 1 misstatement,
using a Normal(0.025, 0.05) prior distribution.

```{r}
prior <- auditPrior(method = "param", likelihood = "normal", alpha = 0.025, beta = 0.05)
eval <- evaluation(materiality = 0.05, x = 1, n = 60, prior = prior)
summary(eval)
```

The prior and posterior distribution can be visualized by calling
`plot(..., type = "posterior")`.

```{r fig.align="center", fig.height=4, fig.width=6}
plot(eval, type = "posterior")
```

Furthermore, the predictions of the posterior distribution at the data level can
be visualized by combining the `predict()` and `plot()` functions. For instance,
the command below displays the predictions of the posterior distribution for a
hypothetical (next) sample of $n$ = 10 items.

```{r fig.align="center", fig.height=4, fig.width=6}
plot(predict(eval$posterior, n = 10))
```

## References
<div id="refs"></div>
