---
title: Digit analysis
author: Koen Derks
output: 
  html_document:
    toc: true
    toc_depth: 3
bibliography: references.bib
csl: apa.csl
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
library(jfa)
```

## Introduction

Welcome to the 'Digit analysis' vignette of the **jfa** package. This page
provides comprehensive examples of how to use the `digit_test()` and
`repeated_test()` functions included in the package.

## Function: `digit_test()`

The `digit_test()` function accepts a vector of numeric values, extracts the
requested digits, and compares the frequencies of these digits to a reference
distribution. By default, the function performs a frequentist hypothesis test of
the null hypothesis that the digits are distributed according to the reference
distribution, and produces a *p*-value. When a prior is specified, the function
performs a Bayesian hypothesis test of the null hypothesis that the digits are
distributed according to the reference distribution against the alternative
hypothesis that the digits are not distributed according to the reference
distribution, and produces a Bayes factor [@kass_1995].

*Practical example:*

Benford's law [@benford_1938] is a principle that describes a pattern in many
naturally-occurring numbers. According to Benford's law, each possible leading
digit $d$ in a naturally occurring, or non-manipulated, set of numbers occurs
with a probability:

\begin{equation}
  p(d_i) = \text{log}_{10}(1 + \frac{1}{d_i})
\end{equation}

The distribution of leading digits in a data set of financial transaction values
(e.g., the `sinoForest` data) can be extracted and tested against the expected
frequencies under Benford's law using the code below.

```{r}
x <- digit_test(sinoForest$value, check = "first", reference = "benford")
print(x)
```

You can visually compare the distribution of first digits to the reference
distribution by calling `plot(..., type = "estimates")` on the returned object.

```{r fig.align="center", fig.height=4, fig.width=6}
plot(x, type = "estimates")
```

You can also conduct this analysis in a Bayesian manner by setting
`prior = TRUE`, or by providing a value for the prior concentration parameter
(e.g., `prior = 3`).

```{r}
x <- digit_test(sinoForest$value, check = "first", reference = "benford", prior = TRUE)
print(x)
```

When performing the analysis in a Bayesian manner, you can invoke
`plot(..., type = "robustness")` to assess the robustness of the Bayes factor to
the choice of the prior distribution. This will display the Bayes factor under
various reasonable specifications of the prior distribution.

```{r fig.align="center", fig.height=4, fig.width=6}
plot(x, type = "robustness")
```

In addition, you can perform a sequential analysis using the Bayes factor by
invoking `plot(..., type = "sequential")`. This sequential analysis includes a
robustness check as well.

```{r fig.align="center", fig.height=4, fig.width=6}
plot(x, type = "sequential")
```

## Function: `repeated_test()`

The `repeated_test()` function analyzes the frequency with which values are
repeated within a set of numbers. Unlike Benford's law, and its generalizations,
this approach examines the entire number at once, not only the first or last
digit. For the technical details of this procedure, see [@simonsohn_2019].

*Practical example:*

In this example, we analyze a data set from a (retracted) paper that describes
three experiments run in Chinese factories, where workers were nudged to use
more hand-sanitizer. These data were shown to exhibit two classic markers of
data tampering: impossibly similar means and the uneven distribution of last
digits [@yu_2018]. We can use the `repeated_test()` function to test if these
data also contain a greater amount of repeated values than expected if the data
were not tampered with.

```{r}
x <- repeated_test(sanitizer$value, check = "lasttwo", samples = 2000)
print(x)
```

A histogram of the frequency of each value can be obtained via the `plot()`
function.

```{r fig.align="center", fig.height=4, fig.width=6}
plot(x)
```

## References
<div id="refs"></div>
